PANCHANGE_data_preprocessing

### 2020-08-19 PANCHANGE_analysis.Rmd
fixed ethnicity categories

Ethnicity should be recoded
European
Mixed or multiplce ethnic origins
Africian or African British
Asian or Asian British
Arab
```{r Disorder by Ethnicity}
dat.long %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Ethnicity) %>%
  count(Disorder)
```

>>>>

Ethnicity should be recoded
European
Mixed or multiple ethnic origins
African or African British
Asian or Asian British
Arab
```{r Disorder by Ethnicity}
dat.long %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Ethnicity) %>%
  count(Disorder)
```

### 2020-08-19 PANCHANGE_data_preprocessing2020-07-15.Rmd
fixed ethnicity categories

glad.dem.raw.id <- glad.dem.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate.prepandemic = startDate, #Date when participant signed up to glad
         Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
         Gender_unc = demographics.which_gender_do_you_identify_with, # Gender
         Sex = demographics.select_questionnaire_items_medical, #Sex
         EduYrs = demographics.school_education_school_years.txt, # Years in education
         demographics.college_or_university_degree, # level of education - univeristy
         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
         demographics.cses_or_equivalent, # level of education - cse
         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
      startDate.prepandemic = startDate, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender_unc = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity (different to GLAD)
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

#Match GLAD/EDGI/RAMP categories
nbr.cohort.raw.id$Ethnicity <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White, white European or Caucasian",
                                             "B" = "White, white European or Caucasian",
                                             "C" = "White, white European or Caucasian",
                                             "D" = "Mixed or multiple ethnic origins",
                                             "E" = "Mixed or multiple ethnic origins",
                                             "F" = "Mixed or multiple ethnic origins",
                                             "G" = "Mixed or multiple ethnic origins",
                                             "H" = "Asian or Asian British (including Chinese)",
                                             "J" = "Asian or Asian British (including Chinese)",
                                             "K" = "Asian or Asian British (including Chinese)",
                                             "L" = "Asian or Asian British (including Chinese)",
                                             "M" = "Black or Black British",
                                             "N" = "Black or Black British",
                                             "P" = "Black or Black British",
                                             "R" = "Asian or Asian British (including Chinese)",
                                             "S" = "Asian or Asian British (including Chinese)",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)
}
```

@@

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(`Login ID`) %>% # Drop NAs
    distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, #Date when participant signed up to ramp  (is in the coping gad chunk)
      age_category_full = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
#     EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
#     demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
#     demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#     demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#     demographics.cses_or_equivalent, # level of education - cse
#     demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r gender recoding for GLAD}
if(GLAD == TRUE) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Gender = 
      recode_factor(Gender_unc,
        "Male" = "Male",
        "Female" = "Female",
        "Non-binary" = "Non-binary",
        "Prefer to self-define (please tell us more)" = "Prefer to self-define"
      )
  )
}
```

@@

dem.raw.id <- dem.raw.id %>%
  mutate(
    Gender = 
      recode_factor(Gender_unc,
        "Male" = "Male",
        "Female" = "Female",
        "Non-binary" = "Non-binary",
        "Prefer to self-define (please tell us more):" = "Prefer to self-define"
      )
  )
}
```

```{r gender recoding for NBR}
if(NBR == TRUE ) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Gender = 
      recode_factor(Gender_unc,
        "Male" = "Male",
        "Female" = "Female",
        "Non-binary" = "Non-binary",
        "Prefer to self-define (please tell us more)" = "Prefer to self-define"
      )
  )
}
```

>>>>

glad.dem.raw.id <- glad.dem.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate.prepandemic = startDate, #Date when participant signed up to glad
         Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
         Gender_unc = demographics.which_gender_do_you_identify_with, # Gender
         Sex = demographics.select_questionnaire_items_medical, #Sex
         EduYrs = demographics.school_education_school_years.txt, # Years in education
         demographics.college_or_university_degree, # level of education - univeristy
         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
         demographics.cses_or_equivalent, # level of education - cse
         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity_unc = demographics.questions_based_ethnic_origin # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
      startDate.prepandemic = startDate, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender_unc = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity_unc = demographics.what_is_your_ethnic_origin # Ethnicity (different to GLAD)
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

#Match GLAD/EDGI/RAMP categories
nbr.cohort.raw.id$Ethnicity_unc <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White, white European or Caucasian",
                                             "B" = "White, white European or Caucasian",
                                             "C" = "White, white European or Caucasian",
                                             "D" = "Mixed or multiple ethnic origins",
                                             "E" = "Mixed or multiple ethnic origins",
                                             "F" = "Mixed or multiple ethnic origins",
                                             "G" = "Mixed or multiple ethnic origins",
                                             "H" = "Asian or Asian British (including Chinese)",
                                             "J" = "Asian or Asian British (including Chinese)",
                                             "K" = "Asian or Asian British (including Chinese)",
                                             "L" = "Asian or Asian British (including Chinese)",
                                             "M" = "Black or Black British",
                                             "N" = "Black or Black British",
                                             "P" = "Black or Black British",
                                             "R" = "Asian or Asian British (including Chinese)",
                                             "S" = "Asian or Asian British (including Chinese)",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)
}
```

@@

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(`Login ID`) %>% # Drop NAs
    distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, #Date when participant signed up to ramp  (is in the coping gad chunk)
      age_category_full = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
#     EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
#     demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
#     demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#     demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#     demographics.cses_or_equivalent, # level of education - cse
#     demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity_unc = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r gender recoding for GLAD}
if(GLAD == TRUE | NBR == TRUE) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Gender = 
      recode_factor(Gender_unc,
        "Male" = "Male",
        "Female" = "Female",
        "Non-binary" = "Non-binary",
        "Prefer to self-define (please tell us more)" = "Prefer to self-define"
      )
  ) %>%
  rename(
    Gender_numeric = Gender_unc_numeric
  )
}
```

@@

dem.raw.id <- dem.raw.id %>%
  mutate(
    Gender = 
      recode_factor(Gender_unc,
        "Male" = "Male",
        "Female" = "Female",
        "Non-binary" = "Non-binary",
        "Prefer to self-define (please tell us more):" = "Prefer to self-define"
      )
  ) %>%
  rename(
    Gender_numeric = Gender_unc_numeric
  )
}
```

# Ethnicity recoding

```{r ethnicity recoding for EDGI}
if(EDGI == TRUE) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Ethnicity = 
      recode_factor(Ethnicity_unc,
        "White" = "European",
        "Mixed" = "Mixed or multiple ethnic origins",
        "Asian or Asian British" = "Asian or Asian British",
        "Black or Black British" = "African or African British",
        "Arab" = "Arab",
        "Other ethnic group:" = "Other ethnic group"
      )
  ) %>%
  rename(
    Ethnicity_numeric = Ethnicity_unc_numeric
  )
}
```

```{r ethnicity recoding for GLAD and RAMP}
if(GLAD == TRUE | RAMP == TRUE) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Ethnicity = 
      recode_factor(Ethnicity_unc,
        "White, white European or Caucasian" = "European",
        "Mixed or multiple ethnic origins" = "Mixed or multiple ethnic origins",
        "Asian or Asian British (including Chinese)" = "Asian or Asian British",
        "Black or Black British" = "African or African British",
        "Arab" = "Arab",
        "Other ethnic group (please specify)" = "Other ethnic group"
      )
  ) %>%
  rename(
    Ethnicity_numeric = Ethnicity_unc_numeric
  )
}
```

```{r ethnicity recoding for NBR}
if(NBR == TRUE) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Ethnicity = 
      recode_factor(Ethnicity_unc,
        "White, white European or Caucasian" = "European",
        "Mixed or multiple ethnic origins" = "Mixed or multiple ethnic origins",
        "Asian or Asian British (including Chinese)" = "Asian or Asian British",
        "Black or Black British" = "African or African British",
        "Arab" = "Arab",
        "Other ethnic group (please specify)" = "Other ethnic group"
      )
  ) %>%
  mutate(
    Ethnicity_numeric = 
      recode(Ethnicity_unc,
        "European" = 1,
        "Mixed or multiple ethnic origins" = 2,
        "Asian or Asian British" = 3,
        "African or African British" = 4,
        "Arab" = 5,
        "Other ethnic group" = 6
      )
  )
}
```
